---
title: "Data Pipeline — Beirne & Sugandi (2023) Replication (Japan)"
author: "FINARTS Group 4"
format:
  html:
    self-contained: true
output-file: index
---

The pipeline initialises with project paths and era map before loading any data.

```{r setup, include=FALSE}
#| message: false
#| warning: false

library(tidyverse)
library(readxl)
library(zoo)
library(lubridate)
library(knitr)
library(kableExtra)

DATA_RAW <- file.path("data", "raw")
DATA_OUT <- file.path("data", "processed")
dir.create(DATA_OUT, showWarnings = FALSE, recursive = TRUE)

# Paper uses 14 Jan 1999 as first working day with all financial data available
START_DATE <- as.Date("1999-01-14")
END_DATE   <- as.Date("2026-06-30")
PAPER_END  <- as.Date("2021-03-31")

# Japanese era -> Gregorian year offset
# Heisei 1 = 1989, Reiwa 1 = 2019
# Era year N = year_start + N - 1 => Heisei: 1988 + N, Reiwa: 2018 + N
era_map <- tibble(
  era_name    = c("平成", "令和"),
  era_offset  = c(1988L, 2018L)
)
```

All six daily financial series share a consistent LSEG CSV format, so a single read function handles them.

```{r load-financials}
#| message: false

# All LSEG daily CSVs share the same 6-column format: Date,Close,High,Low,Open,Volume
read_lseg_daily <- function(file, col_name) {
  read_csv(file,
           col_types = cols(Date = col_date(), Close = col_double(),
                            High = col_double(), Low = col_double(),
                            Open = col_double(), Volume = col_double()),
           show_col_types = FALSE) |>
    select(date = Date, !!col_name := Close) |>
    filter(date >= START_DATE)
}

vix      <- read_lseg_daily(file.path(DATA_RAW, "VIX.csv"), "vix")
nikkei   <- read_lseg_daily(file.path(DATA_RAW, "NIKKEI225.csv"), "nikkei225")
sp500    <- read_lseg_daily(file.path(DATA_RAW, "SP500.csv"), "sp500")
us10y    <- read_lseg_daily(file.path(DATA_RAW, "US10Y.csv"), "us10y")
usdjpy   <- read_lseg_daily(file.path(DATA_RAW, "USDJPY.csv"), "usdjpy")

# JGB yields: skip row 1 (title), row 2 is header
# Date format YYYY/MM/DD, 10Y is column 11 (position 10, 0-indexed)
# Filter to rows where 10Y has valid data (1986-07-05+)
raw_jgb <- read_csv(file.path(DATA_RAW, "jgbcme_all.csv"),
                    skip = 1, show_col_types = FALSE) |>
  rename(date = Date) |>
  mutate(date = ymd(date)) |>
  filter(date >= START_DATE)

jgb_cols <- names(raw_jgb)
# 10Y is at position 10 (0-indexed) = 11th column
jgb10y <- raw_jgb |>
  select(date, jgb10y = all_of(jgb_cols[11])) |>
  mutate(jgb10y = as.numeric(na_if(jgb10y, "-")))
```

Macroeconomic data arrives at mixed frequencies from FRED, BIS, and WUI sources, each requiring different parse logic.

```{r load-macro}
#| message: false

# REER.xlsx — check sheet name dynamically
reer_sheets <- excel_sheets(file.path(DATA_RAW, "REER.xlsx"))
reer_raw <- read_excel(file.path(DATA_RAW, "REER.xlsx"),
                       sheet = reer_sheets[1],
                       col_types = "text")
# First row is a sub-header with NA date and "Close" value; skip it
reer <- reer_raw |>
  filter(!is.na(.data[[names(reer_raw)[1]]])) |>
  rename(date_raw = 1, reer = 2) |>
  mutate(date = as_date(ymd_hms(date_raw)),
         reer = as.numeric(reer)) |>
  select(date, reer) |>
  arrange(date) |>
  filter(date >= START_DATE)

# Japan RGDP (FRED, quarterly)
rgdp <- read_csv(file.path(DATA_RAW, "JAPAN_RGDP.csv"),
                 col_types = cols(Date = col_date(), Value = col_double()),
                 show_col_types = FALSE) |>
  rename(date = Date, rgdp = Value) |>
  filter(date >= START_DATE)

# Japan Nominal GDP (FRED, quarterly, millions of JPY)
nominal_gdp <- read_csv(file.path(DATA_RAW, "JAPAN_NOMINAL_GDP.csv"),
                        col_types = cols(Date = col_date(), Value = col_double()),
                        show_col_types = FALSE) |>
  rename(date = Date, nominal_gdp = Value) |>
  filter(date >= START_DATE)

# World Uncertainty Index (Japan, quarterly)
wui <- read_csv(file.path(DATA_RAW, "WUI_JPN.csv"),
                col_types = cols(observation_date = col_date(), WUIJPN = col_double()),
                show_col_types = FALSE) |>
  rename(date = observation_date, wui = WUIJPN) |>
  filter(date >= START_DATE)
```

The BOJ 6pi-1 CSV uses Shift-JIS encoding and Japanese-era year labels, requiring manual line-by-line parsing before structured extraction.

```{r parse-boj-6pi1}
#| message: false

# BOJ 6pi-1 CSV: Shift-JIS (CP932) encoded, mixed layout
# Use base R's readLines with explicit encoding; readr's locale(encoding="Shift-JIS") crashes
boj_con <- file(file.path(DATA_RAW, "BOJ_6pi-1_portfolio_summary.csv"),
                encoding = "CP932")
boj_lines <- readLines(boj_con, warn = FALSE)
close(boj_con)

# Find monthly figures section
monthly_start <- grep("(Monthly figures)", boj_lines, fixed = TRUE)[1] + 1
monthly_txt <- paste(boj_lines[monthly_start:length(boj_lines)], collapse = "\n")

# Parse as CSV — 29 columns expected
# Liabilities section: cols 15-28 (0-indexed 15-28), repeated layout per month
monthly_raw <- read_csv(I(monthly_txt), col_names = FALSE,
                        col_types = cols(.default = col_character()),
                        guess_max = 500, show_col_types = FALSE)

boj_monthly <- monthly_raw |>
  select(era_col = X1, month_col = X2,
         equity_net = X22, lt_debt_net = X25, st_debt_net = X28) |>
  mutate(
    era_col = na_if(era_col, ""),
    month_col = na_if(month_col, "")
  ) |>
  tidyr::fill(era_col, .direction = "down") |>
  filter(!is.na(month_col)) |>
  mutate(
    month_num = as.integer(str_replace(month_col, "月", "")),
    # Handle 元年 (first year) -> 1年
    era_col_digit = str_replace(era_col, "元", "1"),
    era_type = str_extract(era_col_digit, "^[平成令和]+"),
    era_year = as.integer(str_extract(era_col_digit, "\\d+")),
    year_greg = case_when(
      era_type == "平成" ~ 1988L + era_year,
      era_type == "令和" ~ 2018L + era_year,
      TRUE ~ NA_integer_
    ),
    date = ymd(paste(year_greg, month_num, "01", sep = "-"))
  )

# Parse BOJ numeric values (comma-formatted, "--" = NA)
parse_boj_val <- function(x) {
  x <- str_replace_all(x, ",", "")
  x <- na_if(x, "--")
  as.numeric(x)
}

boj_monthly <- boj_monthly |>
  mutate(
    equity_val = parse_boj_val(equity_net),
    lt_debt    = parse_boj_val(lt_debt_net),
    st_debt    = parse_boj_val(st_debt_net),
    debtsec    = lt_debt + st_debt
  ) |>
  filter(date <= as.Date("2013-12-01"))

debtsec_pre2014 <- boj_monthly |>
  select(date, debtsec)
equity_pre2014 <- boj_monthly |>
  select(date, equity = equity_val)
```

Post-2014 BOJ API files are simpler two-column CSVs read by a shared helper.

```{r load-boj-api}
#| message: false

# Post-2014 BOJ API flows: simple 2-col CSV (date, value), 100M JPY
read_boj_api <- function(file, col_name) {
  read_csv(file,
           col_types = cols(Date = col_date(),
                            .default = col_double()),
           show_col_types = FALSE) |>
    rename(date = Date, !!col_name := 2) |>
    filter(date >= START_DATE)
}

boj_lt   <- read_boj_api(file.path(DATA_RAW, "BOJ_BPPI6E3N5.csv"), "lt_debt")
boj_st   <- read_boj_api(file.path(DATA_RAW, "BOJ_BPPI6E4N5.csv"), "st_debt")
boj_eq   <- read_boj_api(file.path(DATA_RAW, "BOJ_BPPI6E2N5.csv"), "equity")
boj_oth  <- read_boj_api(file.path(DATA_RAW, "BOJ_BPBP6JYNFL3.csv"), "other")
boj_dir  <- read_boj_api(file.path(DATA_RAW, "BOJ_BPBP6JYNFL13.csv"), "direct")
```

Pre-2014 parsed data and post-2014 API data must be concatenated into continuous capital flow series.

```{r merge-capital-flows}
#| message: false

# Post-2014 debtsec = long-term + short-term debt securities (sum per D-04)
debtsec_post2014 <- boj_lt |>
  inner_join(boj_st, by = "date") |>
  mutate(debtsec = lt_debt + st_debt) |>
  select(date, debtsec)

equity_post2014 <- boj_eq |>
  select(date, equity)

# Bind pre-2014 and post-2014
debtsec_all <- bind_rows(debtsec_pre2014, debtsec_post2014) |>
  arrange(date)
equity_all <- bind_rows(equity_pre2014, equity_post2014) |>
  arrange(date)

# Other and Direct are single-source, full range
other_all <- boj_oth |>
  rename(value = other) |>
  select(date, value)
direct_all <- boj_dir |>
  rename(value = direct) |>
  select(date, value)
```

The risk-off binary is constructed from the VIX using a left_join to preserve dates where only US markets are open.

```{r risk-off}
#| message: false

# VIX as anchor (trades every US market day). Left_join keeps dates where VIX
# exists but JGB/Nikkei may not (e.g., Japanese holidays). This preserves risk-off
# episode start dates that fall on Japanese non-trading days (audit C-3).
daily_fin <- vix |>
  left_join(nikkei, by = "date") |>
  left_join(sp500, by = "date") |>
  left_join(us10y, by = "date") |>
  left_join(usdjpy, by = "date") |>
  left_join(jgb10y, by = "date") |>
  arrange(date)

daily_fin <- daily_fin |>
  mutate(
    vix_ma60 = rollmean(vix, k = 60, fill = NA, align = "right"),
    risk_off = if_else(vix >= vix_ma60 + 10, 1L, 0L)
  )
```

The spread between JGB and US Treasury yields is a simple point-in-time subtraction of two daily series.

```{r spread}
#| message: false

daily_fin <- daily_fin |>
  mutate(spread = jgb10y - us10y)
```

Log transformations require positive underlying values, with a guard for WUI observations that can reach zero.

```{r log-transforms}
#| message: false

daily_fin <- daily_fin |>
  mutate(log_nikkei = log(nikkei225))

macro <- reer |>
  left_join(rgdp, by = "date") |>
  left_join(wui, by = "date")

# WUI can be zero in early observations — log would produce -Inf
# Filter to positive values before transform
macro <- macro |>
  mutate(
    log_reer = log(reer),
    log_rgdp = log(rgdp),
    log_wui  = if_else(wui > 0, log(wui), NA_real_)
  )
```

Capital flow %GDP conversion involves end-of-month FX rates and quarterly GDP carried forward to monthly frequency.

```{r capital-flow-pct-gdp}
#| message: false

# End-of-month USD/JPY (last trading day of each month)
usdjpy_eom <- daily_fin |>
  mutate(year_month = floor_date(date, "month")) |>
  group_by(year_month) |>
  filter(date == max(date, na.rm = TRUE)) |>
  ungroup() |>
  distinct(year_month, usdjpy_eom = usdjpy)

# Prepare nominal GDP with year_month key for joining
# GDP is quarterly — expand to full monthly sequence and carry forward
# the quarterly value so months 2-3 of each quarter don't get NA (audit W-4)
nominal_gdp_monthly <- nominal_gdp |>
  mutate(year_month = floor_date(date, "month")) |>
  complete(year_month = seq(min(year_month), max(year_month), by = "month")) |>
  fill(nominal_gdp, .direction = "down")

# Helper: 100M JPY -> USD billion
convert_100m_jpy_to_usd_bn <- function(v, rate) (v * 0.1) / rate

# Helper: millions of JPY -> USD billion
convert_jpy_mn_to_usd_bn <- function(v, rate) v / (rate * 1000)

# Debtsec %GDP
debtsec_pct <- debtsec_all |>
  mutate(year_month = floor_date(date, "month")) |>
  left_join(usdjpy_eom, by = "year_month") |>
  left_join(nominal_gdp_monthly, by = "year_month") |>
  mutate(
    flow_usd_bn   = convert_100m_jpy_to_usd_bn(debtsec, usdjpy_eom),
    gdp_usd_bn    = convert_jpy_mn_to_usd_bn(nominal_gdp, usdjpy_eom),
    debtsec_pct   = (flow_usd_bn / gdp_usd_bn) * 100
  ) |>
  select(date = year_month, debtsec_pct)

# Equity %GDP
equity_pct <- equity_all |>
  mutate(year_month = floor_date(date, "month")) |>
  left_join(usdjpy_eom, by = "year_month") |>
  left_join(nominal_gdp_monthly, by = "year_month") |>
  mutate(
    flow_usd_bn  = convert_100m_jpy_to_usd_bn(equity, usdjpy_eom),
    gdp_usd_bn   = convert_jpy_mn_to_usd_bn(nominal_gdp, usdjpy_eom),
    equity_pct   = (flow_usd_bn / gdp_usd_bn) * 100
  ) |>
  select(date = year_month, equity_pct)

# Other %GDP
other_pct <- other_all |>
  mutate(year_month = floor_date(date, "month")) |>
  left_join(usdjpy_eom, by = "year_month") |>
  left_join(nominal_gdp_monthly, by = "year_month") |>
  mutate(
    flow_usd_bn = convert_100m_jpy_to_usd_bn(value, usdjpy_eom),
    gdp_usd_bn  = convert_jpy_mn_to_usd_bn(nominal_gdp, usdjpy_eom),
    other_pct   = (flow_usd_bn / gdp_usd_bn) * 100
  ) |>
  select(date = year_month, other_pct)

# Direct %GDP
direct_pct <- direct_all |>
  mutate(year_month = floor_date(date, "month")) |>
  left_join(usdjpy_eom, by = "year_month") |>
  left_join(nominal_gdp_monthly, by = "year_month") |>
  mutate(
    flow_usd_bn  = convert_100m_jpy_to_usd_bn(value, usdjpy_eom),
    gdp_usd_bn   = convert_jpy_mn_to_usd_bn(nominal_gdp, usdjpy_eom),
    direct_pct   = (flow_usd_bn / gdp_usd_bn) * 100
  ) |>
  select(date = year_month, direct_pct)
```

Python requires native-frequency data anchored before the sample start, so raw files are re-read without the START_DATE filter.

```{r export-native-data}
#| message: false

# Export native-frequency data (unfiltered) for Python quadratic interpolation.
# Python needs anchor points before 1999-01-14. We re-read raw files without
# the START_DATE filter so the full available history feeds the interpolation.

TMP_DIR <- file.path("data", "tmp_quadratic")
dir.create(TMP_DIR, showWarnings = FALSE, recursive = TRUE)

# Macro data: re-read raw files unfiltered, compute logs
reer_full <- read_excel(file.path(DATA_RAW, "REER.xlsx"),
  sheet = excel_sheets(file.path(DATA_RAW, "REER.xlsx"))[1],
  col_types = "text") |>
  filter(!is.na(.data[[names(reer_raw)[1]]])) |>
  rename(date_raw = 1, reer = 2) |>
  mutate(date = as_date(ymd_hms(date_raw)), reer = as.numeric(reer)) |>
  select(date, reer) |> arrange(date)

rgdp_full <- read_csv(file.path(DATA_RAW, "JAPAN_RGDP.csv"),
  col_types = cols(Date = col_date(), Value = col_double()),
  show_col_types = FALSE) |>
  rename(date = Date, rgdp = Value)

wui_full <- read_csv(file.path(DATA_RAW, "WUI_JPN.csv"),
  col_types = cols(observation_date = col_date(), WUIJPN = col_double()),
  show_col_types = FALSE) |>
  rename(date = observation_date, wui = WUIJPN)

reer_full |> mutate(log_reer = log(reer)) |>
  select(date, log_reer) |>
  write_csv(file.path(TMP_DIR, "log_reer_native.csv"))

rgdp_full |> mutate(log_rgdp = log(rgdp)) |>
  select(date, log_rgdp) |>
  write_csv(file.path(TMP_DIR, "log_rgdp_native.csv"))

wui_full |> filter(wui > 0) |> mutate(log_wui = log(wui)) |>
  select(date, log_wui) |>
  write_csv(file.path(TMP_DIR, "log_wui_native.csv"))

# Capital flow %GDP: re-read BOJ, USDJPY, and Nominal GDP raw files unfiltered
usdjpy_full <- read_csv(file.path(DATA_RAW, "USDJPY.csv"),
  col_types = cols(Date = col_date(), Close = col_double(),
    High = col_double(), Low = col_double(),
    Open = col_double(), Volume = col_double()),
  show_col_types = FALSE) |>
  select(date = Date, usdjpy = Close) |>
  mutate(year_month = floor_date(date, "month")) |>
  group_by(year_month) |>
  filter(date == max(date, na.rm = TRUE)) |>
  ungroup() |>
  distinct(year_month, usdjpy_eom = usdjpy)

ngdp_full <- read_csv(file.path(DATA_RAW, "JAPAN_NOMINAL_GDP.csv"),
  col_types = cols(Date = col_date(), Value = col_double()),
  show_col_types = FALSE) |>
  rename(date = Date, nominal_gdp = Value) |>
  mutate(year_month = floor_date(date, "month")) |>
  complete(year_month = seq(min(year_month), max(year_month), by = "month")) |>
  fill(nominal_gdp, .direction = "down")

to_usd_bn       <- function(v, rate) (v * 0.1) / rate
ngdp_to_usd_bn  <- function(v, rate) v / (rate * 1000)

# Debtsec: combine pre-2014 6pi-1 with post-2014 BOJ API (long + short)
boj_lt_full <- read_csv(file.path(DATA_RAW, "BOJ_BPPI6E3N5.csv"),
  col_types = cols(Date = col_date(), .default = col_double()),
  show_col_types = FALSE) |> rename(date = Date, lt = 2)
boj_st_full <- read_csv(file.path(DATA_RAW, "BOJ_BPPI6E4N5.csv"),
  col_types = cols(Date = col_date(), .default = col_double()),
  show_col_types = FALSE) |> rename(date = Date, st = 2)

debtsec_post_full <- boj_lt_full |>
  inner_join(boj_st_full, by = "date") |>
  mutate(value = lt + st) |> select(date, value)

debtsec_full <- bind_rows(
  debtsec_pre2014 |> rename(value = debtsec), debtsec_post_full
) |> arrange(date)

equity_post_full <- read_csv(file.path(DATA_RAW, "BOJ_BPPI6E2N5.csv"),
  col_types = cols(Date = col_date(), .default = col_double()),
  show_col_types = FALSE) |> rename(date = Date, value = 2)

equity_full <- bind_rows(
  equity_pre2014 |> rename(value = equity), equity_post_full
) |> arrange(date)

# Native-frequency %GDP exports
debtsec_full |>
  mutate(year_month = floor_date(date, "month")) |>
  left_join(usdjpy_full, by = "year_month") |>
  left_join(ngdp_full, by = "year_month") |>
  mutate(debtsec_pct = (to_usd_bn(value, usdjpy_eom) / ngdp_to_usd_bn(nominal_gdp, usdjpy_eom)) * 100) |>
  select(date = year_month, debtsec_pct) |>
  write_csv(file.path(TMP_DIR, "debtsec_native.csv"))

equity_full |>
  mutate(year_month = floor_date(date, "month")) |>
  left_join(usdjpy_full, by = "year_month") |>
  left_join(ngdp_full, by = "year_month") |>
  mutate(equity_pct = (to_usd_bn(value, usdjpy_eom) / ngdp_to_usd_bn(nominal_gdp, usdjpy_eom)) * 100) |>
  select(date = year_month, equity_pct) |>
  write_csv(file.path(TMP_DIR, "equity_native.csv"))

read_csv(file.path(DATA_RAW, "BOJ_BPBP6JYNFL3.csv"),
  col_types = cols(Date = col_date(), .default = col_double()),
  show_col_types = FALSE) |>
  rename(date = Date, value = 2) |>
  mutate(year_month = floor_date(date, "month")) |>
  left_join(usdjpy_full, by = "year_month") |>
  left_join(ngdp_full, by = "year_month") |>
  mutate(other_pct = (to_usd_bn(value, usdjpy_eom) / ngdp_to_usd_bn(nominal_gdp, usdjpy_eom)) * 100) |>
  select(date = year_month, other_pct) |>
  write_csv(file.path(TMP_DIR, "other_native.csv"))

read_csv(file.path(DATA_RAW, "BOJ_BPBP6JYNFL13.csv"),
  col_types = cols(Date = col_date(), .default = col_double()),
  show_col_types = FALSE) |>
  rename(date = Date, value = 2) |>
  mutate(year_month = floor_date(date, "month")) |>
  left_join(usdjpy_full, by = "year_month") |>
  left_join(ngdp_full, by = "year_month") |>
  mutate(direct_pct = (to_usd_bn(value, usdjpy_eom) / ngdp_to_usd_bn(nominal_gdp, usdjpy_eom)) * 100) |>
  select(date = year_month, direct_pct) |>
  write_csv(file.path(TMP_DIR, "direct_native.csv"))
```

Quadratic interpolation runs in Python because R lacks a native quadratic spline; pandas provides it with no extrapolation beyond anchor points.

```{r python-interpolation}
#| message: false

# True quadratic interpolation via Python pandas (matches paper Section 4 method).
# R has no native quadratic spline; pandas provides it as a one-liner with
# proper boundary behaviour (no extrapolation past anchor points).
# Requires: Python 3.10+ with pandas>=2.0, numpy>=1.24 (managed by pyproject.toml)
system2("uv", c("run", "python3", "scripts/quadratic_interpolate.py"))
```

The daily calendar and interpolated series are imported back into R for the working-days filter.

```{r import-daily}
#| message: false

# Full daily calendar (used by working-days filter below)
all_dates <- tibble(date = seq(START_DATE, END_DATE, by = "day"))

# Import quadratic-interpolated daily series from Python output
import_daily <- function(name) {
  read_csv(file.path("data", "tmp_quadratic", paste0(name, "_daily.csv")),
    show_col_types = FALSE)
}

log_reer_daily <- import_daily("log_reer")
log_rgdp_daily <- import_daily("log_rgdp")
log_wui_daily  <- import_daily("log_wui")
debtsec_daily  <- import_daily("debtsec")
equity_daily   <- import_daily("equity")
other_daily    <- import_daily("other")
direct_daily   <- import_daily("direct")
```

Weekends are removed after interpolation, keeping only Monday-through-Friday trading days.

```{r working-days}
#| message: false

# Remove weekends; keep Mon-Fri trading days
is_working_day <- function(d) wday(d) %in% 2:6

working_dates <- all_dates |>
  filter(is_working_day(date))

filter_working <- function(df) {
  df |> filter(date %in% working_dates$date)
}

risk_off_daily   <- daily_fin |> select(date, risk_off)
spread_daily     <- daily_fin |> select(date, spread)
log_nikkei_daily <- daily_fin |> select(date, log_nikkei)

risk_off_wd   <- filter_working(risk_off_daily)
spread_wd     <- filter_working(spread_daily)
log_nikkei_wd <- filter_working(log_nikkei_daily)
log_reer_wd   <- filter_working(log_reer_daily)
log_rgdp_wd   <- filter_working(log_rgdp_daily)
log_wui_wd    <- filter_working(log_wui_daily)
debtsec_wd    <- filter_working(debtsec_daily)
equity_wd     <- filter_working(equity_daily)
other_wd      <- filter_working(other_daily)
direct_wd     <- filter_working(direct_daily)
```

All ten variables are joined on date in Cholesky order and assigned a period flag at the paper's sample boundary.

```{r merge-all}
#| message: false

final_dataset <- risk_off_wd |>
  full_join(log_wui_wd, by = "date") |>
  full_join(spread_wd, by = "date") |>
  full_join(log_rgdp_wd, by = "date") |>
  full_join(log_reer_wd, by = "date") |>
  full_join(log_nikkei_wd, by = "date") |>
  full_join(debtsec_wd, by = "date") |>
  full_join(equity_wd, by = "date") |>
  full_join(other_wd, by = "date") |>
  full_join(direct_wd, by = "date") |>
  arrange(date) |>
  mutate(
    period = case_when(
      date <= PAPER_END ~ "1999–2021 (Paper Period)",
      TRUE               ~ "2021–2026 (Extended)"
    )
  )
```

The final merged dataset is written to disk as a single CSV.

```{r export-csv}
#| message: false

write_csv(final_dataset, file.path(DATA_OUT, "final_dataset.csv"))
cat("final_dataset.csv written:", nrow(final_dataset), "rows x", ncol(final_dataset), "cols\n")
```

Descriptive statistics are computed per period and exported both as a rendered table and a CSV file.

```{r descriptive-stats}
#| message: false

desc_period <- final_dataset |>
  select(-date) |>
  pivot_longer(-period, names_to = "variable", values_to = "value") |>
  group_by(period, variable) |>
  summarise(
    N    = sum(!is.na(value)),
    Mean = mean(value, na.rm = TRUE),
    SD   = sd(value, na.rm = TRUE),
    Min  = min(value, na.rm = TRUE),
    Max  = max(value, na.rm = TRUE),
    .groups = "drop"
  )

desc_total <- final_dataset |>
  select(-date, -period) |>
  pivot_longer(everything(), names_to = "variable", values_to = "value") |>
  group_by(variable) |>
  summarise(
    N    = sum(!is.na(value)),
    Mean = mean(value, na.rm = TRUE),
    SD   = sd(value, na.rm = TRUE),
    Min  = min(value, na.rm = TRUE),
    Max  = max(value, na.rm = TRUE),
    .groups = "drop"
  )

desc_table <- desc_period |>
  pivot_longer(c(N, Mean, SD, Min, Max), names_to = "stat") |>
  pivot_wider(names_from = period, values_from = value) |>
  left_join(
    desc_total |> pivot_longer(-variable, names_to = "stat"),
    by = c("variable", "stat")
  ) |>
  rename(`Full Sample` = value) |>
  arrange(variable, stat)

desc_table |>
  mutate(across(where(is.numeric), ~ round(.x, 4))) |>
  select(-variable) |>
  kbl(caption = "Descriptive Statistics by Period", col.names = c("Stat",
    "1999–2021 (Paper Period)", "2021–2026 (Extended)", "Full Sample")) |>
  kable_styling(bootstrap_options = c("hover"), font_size = 11) |>
  pack_rows(index = table(factor(desc_table$variable,
    levels = unique(desc_table$variable)))) |>
  column_spec(1, width = "4em") |>
  column_spec(2, background = "#D6DBF5") |>
  column_spec(3, background = "#FEE1CD") |>
  column_spec(4, background = "#F2F2F2")

write_csv(desc_table, file.path(DATA_OUT, "descriptive_stats.csv"))
cat("descriptive_stats.csv written:", nrow(desc_table), "rows\n")
```

## Visualizing the Descriptive Statistics

Three figures help diagnose distributional shifts, structural breaks and cross-variable relationships in the extended sample.

```{r viz-setup}
#| message: false

# QVRS-compliant theme for all ggplot2 figures
theme_quant <- function(base_size = 10) {
  theme_minimal(base_size = base_size) +
    theme(
      text = element_text(color = "#0D0D0D"),
      plot.title = element_text(face = "bold", size = rel(1.3), margin = margin(b = 5)),
      plot.subtitle = element_text(size = rel(1), color = "#333333", margin = margin(b = 8)),
      plot.caption = element_text(size = rel(0.8), color = "#595959", hjust = 0, margin = margin(t = 8)),
      axis.title = element_text(face = "bold", size = rel(0.9)),
      axis.text = element_text(size = rel(0.85)),
      axis.ticks = element_blank(),
      panel.grid.minor = element_blank(),
      panel.grid.major = element_line(color = "#F2F2F2", linewidth = 0.3),
      strip.text = element_text(face = "bold", size = rel(0.9)),
      legend.position = "bottom",
      legend.text = element_text(size = rel(0.85))
    )
}

# QVRS palette: Chicago 45 (paper), Singapore 55 (extended), London 35 (full sample)
period_colors <- c("#2E45B8", "#F97A1F", "#595959")
names(period_colors) <- c("1999–2021 (Paper Period)", "2021–2026 (Extended)", "Full Sample")
```

### Distributional shifts after 2021

Overlaid densities show how each variable\'s distribution changed between the paper period and the extended sample. The vertical dashed line marks the full-sample mean.

```{r viz-density}
#| fig.height: 10
#| fig.width: 10
#| message: false

plot_data <- final_dataset |>
  select(-date) |>
  pivot_longer(-period, names_to = "variable", values_to = "value") |>
  filter(!is.na(value))

plot_data$period <- factor(plot_data$period,
  levels = c("1999–2021 (Paper Period)", "2021–2026 (Extended)", "Full Sample"))

full_means <- plot_data |>
  filter(period == "Full Sample") |>
  group_by(variable) |>
  summarise(full_mean = mean(value, na.rm = TRUE), .groups = "drop")

plot_data <- plot_data |> left_join(full_means, by = "variable")

ggplot(plot_data, aes(x = value, fill = period, color = period)) +
  geom_density(alpha = 0.25, linewidth = 0.5) +
  geom_vline(data = full_means, aes(xintercept = full_mean),
    color = "#0D0D0D", linewidth = 0.7, linetype = "dashed") +
  facet_wrap(~variable, scales = "free", ncol = 3) +
  scale_fill_manual(values = period_colors) +
  scale_color_manual(values = period_colors) +
  labs(x = "Value", y = "Density", fill = NULL, color = NULL,
       title = "Japanese financial variables show wider dispersion after 2021",
       caption = "Dashed line: full-sample mean. Source: Authors\' calculations using Bloomberg, BOJ, BIS, FRED, IMF data.") +
  theme_quant()
```

### Structural breaks in the extended sample

Each variable\'s daily series from 1999 to 2026, shaded where the extended period begins. The vertical line marks 31 March 2021, the end of the paper\'s estimation window.

```{r viz-timeseries}
#| fig.height: 12
#| fig.width: 10
#| message: false

ts_data <- final_dataset |>
  select(date, period, risk_off, spread, log_nikkei, log_reer, log_rgdp,
         log_wui, debtsec_pct, equity_pct, other_pct, direct_pct) |>
  pivot_longer(-c(date, period), names_to = "variable", values_to = "value") |>
  filter(!is.na(value))

paper_end <- as.Date("2021-03-31")

ggplot(ts_data, aes(x = date, y = value)) +
  geom_line(aes(color = period), linewidth = 0.25, alpha = 0.7) +
  geom_vline(xintercept = paper_end, linetype = "dashed",
             color = "#595959", linewidth = 0.3) +
  facet_wrap(~variable, scales = "free_y", ncol = 2) +
  scale_color_manual(values = period_colors) +
  labs(x = "Year", y = NULL, color = NULL,
       title = "Capital flow and macro series reveal regime changes after 2021",
       caption = "Vertical line: paper period end (31 March 2021). Source: Authors\' calculations.") +
  theme_quant() + coord_cartesian(xlim = c(as.Date("1999-01-01"), as.Date("2026-06-30")))
```

### Cross-variable correlations

Pairwise correlations for the full sample, ordered by strength and direction. Positive correlations point toward the Chicago blue end of the scale and negative toward the Tokyo red end.

```{r viz-correlation}
#| fig.height: 7
#| fig.width: 8
#| message: false

library(corrplot)

cor_data <- final_dataset |>
  select(-date, -period)

cor_matrix <- cor(cor_data, use = "pairwise.complete.obs")

corrplot(cor_matrix, method = "color", type = "upper",
  tl.col = "#0D0D0D", tl.cex = 0.65,
  col = colorRampPalette(c("#C91D42", "#FFFFFF", "#2E45B8"))(200),
  addCoef.col = "#0D0D0D", number.cex = 0.5,
  cl.cex = 0.55, mar = c(1, 0, 2, 0),
  title = "Correlation matrix for the full sample, 1999 to 2026")
```
